  
\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{tcolorbox}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[
top    = 2.50cm,
bottom = 2.50cm,
left   = 2.75cm,
right  = 2.75cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{Analysis 1}
\rhead{EPFL/Alp Ozen}

\newtheorem{remark}{Remark}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{prop}{[Proposition]}
\newtheorem{definition}{Definition}
\numberwithin{equation}{subsection}

\title{Analysis}
\author{SemesterOne analysis at EPFL}
\date{\vspace{-5ex}}
\newtheorem{example}{Example}[section]
\newtheorem{axiom}{Axiom}
\newtheorem{cor}{Corollary}

\begin{document}

\maketitle

\section{Proofs and the reals}
\subsection{Some general proofs}
A valid proof is set of lines where each line logically follows from the next. 
A most famous proof is that $\sqrt{2} $ is irrational. 

\begin{tcolorbox}
\begin{proof}
	Suppose that $\sqrt{2} = \frac{a}{b}$ where $a,b \in \mathbb{Z}$ and $gcd(a,b) = 1$
	\\
	Now we have that $\sqrt{2}b = a$ which means that $2b^2 = a^2$.
	As result, $2 \vert a^2$ hence also $2 \vert a$. Thus we get that $a = 2k$ which also means that $b^2 = 2k^2$ hence $2 \vert b$. As result, $gcd(a,b) = 2$ which is a contradiction. Therefore, $\sqrt{2}$ must be irrational. 
	\end{proof}
\end{tcolorbox}

Quite interestingly, we can also construct a 'wrong' proof just through one fallacious assumption and a set of correct steps. 

\begin{tcolorbox}
    \textbf{Claim:} 1 is the largest integer.
    \\
    Proof: 
    \\
    Let $n$ be the largest integer. Then we have $n \geq n^2$. Which also means $0 \geq n^2 - n = n(n-1)$. Now we have that either $n < 0$ or $n - 1 < 0$. But we know that $ n \nless 0$ as $n$ is at least 1. Hence, $n - 1 < 0$ giving us the result $n < 1$ proving our theorem. Note that the mistake here is solely the assumption we made at the start that there was a largest integer. 
\end{tcolorbox}
\subsection{Proofs relating to infinite processes}
Consider the claim that $0.999\ldots = 1$.
One way to prove this claim, rather naively is this.

\begin{align*}
    9 \times 0.999\ldots \\
    = (10 - 1) \times 0.999\ldots\\
    = 9.999\ldots - 0.999\ldots 
    = 1
\end{align*}

Now a more formal proof is to use an infinite sum and limits. Here it is. 

\begin{tcolorbox}
\textbf{Analysis proof of 0.999\ldots = 1}
    \begin{align*}
        0.999\ldots = 9\lim_{k\to\infty}\sum_{i=1}^{k} (10^{-k})\\
        \lim_{k\to\infty}\sum_{i=1}^{k}(10^{-k}) = \frac{10^{-1} - 10^{-(k+1)}}{1-10^{-1}}\\
        = 9 \times \frac{1}{10} \times \frac{10}{9}\\
        = 1
    \end{align*}
\end{tcolorbox}

\subsection{Basic notions of sets}
The breakdown of sets used in 'standard' analysis are $\mathbb{N} \subseteq \mathbb{Z} \subseteq \mathbb{Q} \subseteq \mathbb{R}$.
\\
There are also some common set related notation that must be known. 
\begin{itemize}
    \item a \textbf{subset} $a \subseteq b$ is defined as $\{x \in b \vert \text{"condition"}\}$
    \item a \textbf{open interval} is defined as $\left]a,b[; r\in A, a < r < b$
    \item an \textbf{open ball} $B(a,\lambda) = \left]a-\lambda,a+\lambda[$
\end{itemize}

\subsection{The Reals}

The reals, denoted \mathbb{R} are an ordered field. Here is a more precise definition.

\begin{tcolorbox}
The reals are a set that have the 3 following axioms:
\begin{itemize}
    \item \mathbb{R} is an abelian group under (+) and \mathbb{R*} is an abelian group under ($\times$). In addition to this, multiplication distributes over addition.
    \item The order relation $\leq$ holds $\forall x \in \mathbb{R}$ That is:
    \begin{align*}
        x \leq y \otimes y \leq x\\
        x \leq y, y \leq x \implies x = y\\
        x \leq y \implies \forall a \in \mathbb{R},\  x + a \leq y + a\\
        0 \leq x, 0 \leq y \implies 0 \leq xy
    \end{align*}
    \item The inf and sup axioms hold
\end{itemize}

We shall now come the \textbf{inf} and \textbf{sup} axioms. It should be intuitively clear that any subset of \mathbb{R} 
\end{tcolorbox}

\subsection{Bounds}
Take some subset S in \mathbb{R}. An element B is called an upper bound of S if $\forall x \in S, B \geq x$. Similarly, it is a lower bound of S if $\forall x \in S, B \leq x$.
\\
The maximum B of a set S denoted $max(S)$ is such that $B \in S, \forall x \in S B \geq x$. 
\\
The supremum of a set S(if it exists) is the lowest upper bound. That is $sup(S) = b$ is such that,


    \begin{align}
        \forall x \in S, b \geq x\\
        \forall \epsilon > 0, \exists x_{\epsilon}, b - x_{\epsilon} \leq \epsilon
    \end{align}
    
    
    
    
    
\begin{remark}

   In our above definition, b does not have to be in S.

\end{remark}


\begin{remark}
    Condition 1 states that b is an upper bound of S. 
 
\end{remark}


\begin{remark}
       Given condition 1, b is the \textit{minimum} of the upper bounds of S.
\end{remark}
    
Some examples of $sup$ and $inf$

\begin{example}
\begin{align*}
 Sup ]a,b[ = b\\
 Inf ]a,b[ = a\\
 Sup \{ x\in \mathbb{R} \vert x = 2k\} \implies \text{Sup doesn't exist.}
\end{align*}
\end{example}


We now establish the infinimum axiom.
\begin{axiom}
All non-empty subsets of $\mathbb{R}_{+}^{*}$ have a highest lower boundary(aka.infinimum)
\end{axiom}

\subsection{\mathbb{Q} is dense in \mathbb{R}}
We claim that between every real number, one can find a rational number. Here's the proof,

\begin{tcolorbox}
\begin{proof}
    Let $x<y \in \mathbb{R}$ Suppose now that $\exists a \in \mathbb{Q}$ such that $x<a<y$. By the Archimedean principle(there is always a greater natural number, $n > \frac{1}{y-x}$ which implies $ny>nx+1$ Now since $ny>nx+1$, there is guaranteed to be some integer in the open bound $]nx,ny[$ which we denote $P$. Dividing by $n$, we get that $\frac{P}{n} \in ]x,y[$ which proves the theorem. 
    \tag*{\qedhere}
\end{proof}
\end{tcolorbox}

\subsection{Integer and fractional part}

Any number $\in \mathbb{R}$ has a integer and fractional part(at least intuitively). Let's formally define these. For some $x \in \mathbb{R}$, let $S := \{n\in\mathbb{N} \vert n > x\}$ Now since $S$ is bounded from below, letting $N$ be the minimum of this set, we obtain that $N \not\in S$. N-1 is thus called the integer part of x denoted [x]. ie. [6.4] = 6
\\
Similarly, the fractional part of x denoted {x} is simply ${x} = x - [x]$

\subsection{Pinning it down:Sup/Inf, bounds, max/min}
\begin{definition}
for a given set $S \subseteq \mathbb{R}$, we have the following:
\\
\textbf{Sup s = b} $\iff \forall \epsilon > 0, \exists x_{\epsilon} \in S, s.t. b - x_{\epsilon} < \epsilon$ (resp. Inf s has the flipped argument)
\\
\textbf{Upper bound = b} $\iff \forall x \in S, b\geq x$(resp. lower bound)
\\
\textbf{Max s = b} $\iff$ b is an upper bound and $b \in S$

\end{definition}
\\

And for the sake of repeating the early axiom(but very important) the infimum axiom is:
\begin{axiom}
For all non-empty subsets of $\mathbb{R}$, $inf S$ exists. 
\end{axiom}

Now we make the first claim in this course that uses an epsilon proof.
\begin{tcolorbox}

\begin{prop}
Whenever $S \subseteq \mathbb{N}$, then $inf S = min S$
\end{prop}

\begin{proof}
    Now, by our axiom, we have that $\mathbb{N} \subseteq \mathbb{R}$ hence we know that $inf S$ exists. We now have to show that $inf S = min S$. 
    \\
    
    Suppose that $inf S \not =  min S$ and let $inf S = b$. Now clearly, $b+\epsilon$ is not a lower bound of S. Now, let $\epsilon = \frac{1}{2}$. Because, $b + \epsilon$ is not a lower bound, we know that $\exists s_{e} < b + \epsilon$. 
    \\
    
    Now $s_{e$ is also not a lower bound, so let's pick $\epsilon^{\prime\prime} = s_{\epsilon} - d$. Now again, $s_{\epsilon^{\prime\prime}}$ must exist. We obtain yet the following:
    \\
    \begin{equation*}
        d < s_{\epsilon^{\prime\prime}} < s_{\epsilon} < d + 1/2 
    \end{equation*}
    
    Now, two natural numbers clearly can not be in an interval which is only $\frac{1}{2}$ units long. Hence, contradiction which means that $d \in S$
    \tag*{\qedhere}
\end{proof}

\end{tcolorbox}
\\
Let's now prove that $\sqrt{2}$ belongs to the reals. For this, we need the following corollary and axiom. 

\begin{cor}
Every non-empty subset of \mathbb{R} with an upper boundary admits a supremum. 
\end{cor}

\begin{axiom}
An ordered field F, which \mathbb{R} is, has the Archimedean property if given any positive x and y in F, $\exists n \in \mathbb{Z}$ s.t. $nx > y$
\end{axiom}

\begin{tcolorbox}

\begin{prop}
$\sqrt{2} \in \mathbb{R}$
\end{prop}

\begin{proof}
Suppose we define a set $S = \{r \in \mathbb{R} \vert r \geq 0, r^{2} < 2\}$. Now, as $S$ is a non-empty subset of $\mathbb{R}$ bounded from above(i.e. 2 is an upper bound) we know that $sup S = x$ exists. Our goal is to show that both $x^{2} < 2$ and $x^{2} > 2$ lead to a contradiction. 
\\

\textbf{Case 1:} Suppose $x^{2} < 2$. We want to find $x + \frac{1}{2} \in S$ which implies that $x$ is not an upper bound as $x < x + \frac{1}{2}$ 
\\
$$(x+\frac{1}{2})^2 = x^2 \frac{2}{x} + \frac{1}{n^2} \leq x^2 + \frac{2}{x} + \frac{1}{n} = x^2 + \frac{1}{n}(2x + 1) $$
\\
Now, we want to show that we can pick an $n$ s.t. $ x^2 + \frac{1}{n}(2x + 1) < 2$. If we can pick such an $n$, then we know by transitivity of $<$ that $(x+\frac{1}{n})^2 < 2$ as  $x^2 \frac{2}{x} + \frac{1}{n^2} \leq x^2 + \frac{2}{x} + \frac{1}{n}$
\\
Reordering the terms, we get $\frac{1}{n} < \frac{2 - x^2}{2x + 1}$ and clearly, $\frac{2 - x^2}{2x + 1}$ is positive as $x^2 < 2$ and $x \geq 0$. This way, we apply the archimedean property to know that $n$ exists s.t. $\frac{1}{n} < \frac{2 - x^2}{2x + 1}$. Given this, we now know that  $ x^2 + \frac{1}{n}(2x + 1) < 2$ which in turn implies $x + \frac{1}{n} \in S$ This contradicts that $x = sup S$ hence $x^2 \not < 2$
\\

\textbf{Case 2:} In turn, we consider the case where $x^2 > 2$ and try to derive a contradiction. We want to show that $\exists m \in \mathbb{N}$ s.t. $x - \frac{1}{m}$ is also an upper bound of $S$ which would mean that $x \not = inf S$. Now:

$$ (x - \frac{1}{m})^2 = x^2 - \frac{2x}{m} + \frac{1}{m^2} > x^2 - \frac{2x}{m}$$

We want to choose $m$ s.t. $x^2 - \frac{2x}{m} > 2$. This way, if $x^2 - \frac{2x}{m} > 2$ holds, then since $(x - \frac{1}{m})^2 > x^2 - \frac{2x}{m}$, we will have that $x - \frac{1}{m}$ is an upper bound. We obtain:
$$\frac{x^2 -2}{2x} > \frac{1}{m}$$
\\
Now as $\frac{x^2 -2}{2x}$ is positive, ${1}{m}$ dpes exist. Hence, $x - \frac{1}{m}$ is also an upper bound that implies $x \not = sup S$ if $x^2 > 2$.
\\
Therefore, $sup S = x = \sqrt{2}$ and since every $sup S \in \mathbb{R}$, $\sqrt{2} \in \mathbb{R}$
   
    \tag*{\qedhere}
\end{proof}
\end{tcolorbox}

\subsection{More theorems about \mathbb{R}}

\begin{tcolorbox}
\begin{prop}
If $a < b$ are real numbers, then $\exists c \in \mathbb{Q}, a < c < b$ 
\end{prop}

\begin{proof}
Now our goal is to show that for any real number $a,b$ we can always find such a $c$. 
Now take some arbitrary $n$ and set it to $n = [ \frac{1}{b-a}] + 1$ Now clearly, $n>\frac{1}{b-a}$ and hence $\frac{1}{n} < b-a$. We will now use this result. Realize that $a = \frac{an}{n} < \frac{[an + 1]}{n} \leq \frac{an + 1}{n} = a + \frac{1}{n} < a + b - a = b$ Therefore, we have found that $a < \frac{[an]+1}{n}< b$ where $c = \frac{[an]+1}$
   
    \tag*{\qedhere}
\end{proof}
\end{tcolorbox}

\begin{tcolorbox}
\begin{prop}
If $a < b$ are real numbers, then $\exists c \in \mathbb{R\backslash Q}, a < c < b$ 
\end{prop}

\begin{proof}
Using the above proposition, we know that $\exists c$ for any $\frac{a}{\sqrt{2}} < \frac{b}{\sqrt{2}} $ and now we get that $a < \sqrt{2}c < b $ where $\sqrt{2}c$ is irrational as whenever one term in $x\times y = z$ is rational and the other irrational, we have (supposing $x$ is irrational) $x = \frac{z}{y}$ and if also $z$ were rational, it would make $x$ rational which is a contradiction. 
    \tag*{\qedhere}
\end{proof}
\end{tcolorbox}

\begin{definition}
The absolute value function $x$ is a function $f: \mathbb{R} \to \mathbb{R^{+}}$ such that:

\[ \begin{cases}  
      f(x) = x , x \geq 0 \\
      f(x) = - x , x < 0
   \end{cases}
\]
\end{definition}

Absolute value respects multiplication and division, that is:

\begin{align*}
    |a||b| = |ab|\\
    \frac{|a|}{|b|} = |\frac{a}{b}|
\end{align*}

But this doesn't hold for addition. For addition we have the triangle inequality: 

\begin{equation*}
    |x+y| \leq |x| + |y| 
\end{equation*}

To prove the above: 


\begin{proof}
Take $|x+y| < 0$. Then $|x+y| = -(x+y) = -x -y \leq |x| + |y|$ Take $|x+y| \geq 0$. Then $|x+y| = x + y \leq |x| + |y|$
    \tag*{\qedhere}
\end{proof}
\clearpage
\section{Sequences}

\subsection{Basics}
\begin{tcolorbox}
Let's begin by formally defining sequences. A sequence is a function $f:\mathbb{N} \to \mathbb{R}$ generally denoted $(x_{n})_{n\geq0}$ And here are some more definitions on sequences:
\\
A sequence is:
\begin{itemize}
    \item constant if $\exists \ C \in \mathbb{R};\ x_{n} = C \ \forall \ n \in \mathbb{N}$
    \item bounded from below(resp. above) if $\exists \ m \in \mathbb{R} \ ; m \leq x_{n} \forall \ n \in \mathbb{N}$
    \item bounded if bounded from both directions
    \item increasing(resp. decreasing) if $x_{n+1} \geq x_{n} \forall \ n \in \mathbb{N}$
    \item strictly increasing(resp.decreasing) if $x_{n+1} \leq x_{n} \forall \ n \in \mathbb{N}$
    \item monotonous if it is increasing or decreasing(resp. strictly)
\end{itemize}
\end{tcolorbox}

Let's now consider a proof on the following proposition:

\begin{tcolorbox}
Define $x_{n} = \sqrt{4 + x_{n-1}}, \ x_{0} = 1$. We claim that $x_{n}$ is bounded and more precisely that $1 \leq x \leq 3$.

\begin{proof}
Base case: $x_{0} = 1$ hence holds.
\\
Now supposing proposition is true for all $n-1$ we get

$$ 1 \leq x_{n-1} \leq 3 $$
$$ 5 \leq x_{n-1} + 4 \leq 7$$

Now we get:

$$ \sqrt{5} \leq \sqrt{x_{n-1} + 4} \leq \sqrt{7}$$
$$ 1 \leq \sqrt{5} \leq \sqrt{x_{n-1} + 4} \leq \sqrt{7} \leq 3 $$
\tag*{\qedhere}
\end{proof}
\end{tcolorbox}


\begin{definition}
A sequence $x_{n} $\mathbb{conveges} to $x$ if $\forall \epsilon > 0$ we can find $n_{0} \in \mathbb{N}$ such that $n\geq n_{0} \rightarrow |x_{n} - x| \leq \epsilon $
\end{definition}

Now in an intuitive sense, suppose the sequence converges to $x$ from both the right and the left. $x$ being our central point, we move a distance $\epsilon$ away from this $x$. Now, if we can pick some $n_{0}$ such that for another $n \geq n_{0}$ we have that $|x_{n} - x| \leq \epsilon$ this means that no matter how small we make epsilon we are able to find some $x_{n}$ in this region. 
\\
Having established convergence, any sequence that is not convergent is said to be \textbf{divergent}. 
\\
Let's now prove that a sequence is divergent. 

\begin{tcolorbox}

\begin{proof}

Take $x_{n} = (-1)^{n}$ Now suppose that $x_{n}$ converges to $x$. Then for $\epsilon = \frac{1}{2}, \ \exists n_{\frac{1}{2}} \in \mathbb{N}$ s.t. $\forall n \geq n_{\frac{1}{2}}$ we would have $|x_{n} - x| \leq \frac{1}{2}$. In particular if $n^{\prime}$ is any other integer $n^{\prime}>n_{\frac{1}{2}}$ then we would have $x_{n} - x_{n^{\prime}} \leq |x_{n} - x| + |x - x_{n^{\prime}}| \leq \frac{1}{2} + \frac{1}{2}$ which implies a contradiction as $|x_{n} - x_{n+1}| = 2 > 1$
\tag*{\qedhere}
\end{proof}
\end{tcolorbox}

\subsection{Limits and their algebra}
\begin{definition}
If a sequence $x_{n}$ converges to some $x$, we say that $x$ is the \textbf{limit} of the sequence and is denoted $\lim_{n\to \infty}{x_{n}} = x $ 
\end{definition}

Here are some properties of limits:

\begin{tcolorbox}
For sequences $x_{n}$ and $y_{n}$ with limits $x,y$ we have:
\begin{itemize}
    \item $$\lim_{n \to \infty} {x_{n} \cdot y_{n}} =  x\cdot y $$
    \item $$\lim_{n \to \infty} {x_{n} + y_{n}} =  x+y $$
    \item $$\lim_{n \to \infty} \frac{x_{n}}{y_{n}} = \frac{x}{y} ; y \not = 0$$
    \item if $\exists \ n_{0} \in \mathbb{N}, \ x_{n} \leq y_{n} \ \forall n \geq n_{0}$ then $x\leq y$ 
\end{itemize}
\end{tcolorbox}

We now introduce the famous \textbf{squeeze theorem} and prove it. 

\begin{theorem}
Squeeze theorem
\\
Let $a_{n}$ and $b_{n}$ be both sequences that converge to $a$. In addition, let $c_{n}$ be such that $ \exists n_{0} \in \mathbb{N}, \ \forall n \geq n_{0}, \ a_{n} \leq c_{n} \leq b_{n}$ Then we clearly have the following:
\begin{equation}
\label{1}
\forall \epsilon > 0, \ \exists N_{1}, \ n \geq N_{1} \rightarrow |a_{n} - a| < \epsilon \equiv a - \epsilon < a_{n} < a + \epsilon
\end{equation}
Similarly, for $b_{n}$ we have that:
\begin{equation}
\label{2}
     \forall \epsilon > 0, \ \exists N_{2}, \ n \geq N_{2} \rightarrow |b_{n} - a| < \epsilon \equiv b - \epsilon < b_{n} < a + \epsilon
\end{equation}
Now set $N = \max{\{N_{1},N_{2},n_{0}\}}$. Now since $N \geq N_{1},N_{2},n_{0}$ we have that both \ref{1} and \ref{2} hold $\forall n > N$ This further gives us the result that:
\begin{proof}
\begin{align*}
    a - \epsilon < a_{n} \leq c_{n} \leq b_{N} < a + \epsilon \\
    |c_{n} - a| < \epsilon
\end{align*}
\tag*{\qedhere}
\end{proof}


\end{theorem}
























\end{document} 
